---
title: "omics-tools-table"
author: "María José Romero Lado"
format: html
editor: visual
---

This repository contains a table that summarizes some of the available tools for analyzing omics data. 

```{r setup, include=FALSE}
# Load required libraries
library(DT)
```

```{r, echo=FALSE}
# Create a data frame for the tools
genomics_tools <- data.frame(
  Tool = c("FastQC", "Bowtie", "IGV", "Trimmomatic"),
  Category = c("Analysis", "Sequencing", "Visualization", "Analysis"),
  Description = c(
    "Quality control for sequencing data",
    "Fast and memory-efficient aligner",
    "Interactive genome viewer",
    "Trimming and filtering of reads"
  ),
  Link = c(
    "<a href='https://www.bioinformatics.babraham.ac.uk/projects/fastqc/' target='_blank'>Visit</a>",
    "<a href='http://bowtie-bio.sourceforge.net/index.shtml' target='_blank'>Visit</a>",
    "<a href='https://software.broadinstitute.org/software/igv/' target='_blank'>Visit</a>",
    "<a href='http://www.usadellab.org/cms/?page=trimmomatic' target='_blank'>Visit</a>"
  ),
  Comments = c(
    "Great for identifying issues in raw reads",
    "Often used in RNA-seq pipelines",
    "Supports various genome formats",
    "Handles both single-end and paired-end data"
  ),
  stringsAsFactors = FALSE
)

# Render the table
datatable(
  genomics_tools,
  escape = FALSE, # Allow HTML in the Link column
  options = list(
    pageLength = 5,
    autoWidth = TRUE
  )
)
```

### 🆕 How to Add New Tools to the Table

To add new tools to the table, follow these steps:

1. Open the R Markdown (`.Rmd`) or Quarto (`.qmd`) file in RStudio or a text editor.
2. Locate the code chunk labeled `tools-table`. It contains the `genomics_tools` data frame that defines the table contents.
3. Add a new row to the `data.frame` in the following format:

```{r}
# Example: Adding a new tool
new_row <- data.frame(
  Tool = "New Tool Name",
  Category = "Tool Category (e.g., Sequencing)",
  Description = "A short description of the tool",
  Link = "<a href='URL' target='_blank'>Visit</a>",
  Comments = "Any relevant comments about the tool",
  stringsAsFactors = FALSE
)

# Append the new row to the existing table
genomics_tools <- rbind(genomics_tools, new_row)
```

